Jager: Symbolic Specification-Based Program Adjustment

### 2013-04-30 paper/paper.tex
Wrapping some quoted words in italics instead.
All the sections were one level too low, I have promoted them.

\section{Outline}

\subsection{Introduction}
The field of program synthesis has seen dramatic improvements over the last
few years.  The increasing power of SAT solvers has allowed researchers to
ask more sophisticated questions and to handle larger programs than they
ever have before.  In short, this is an exciting time in program synthesis.
With the increased power available, the field has progressed beyond the
merely theoretical stage, to one that is making real contributions with even
larger gains expected in the future.

@@

\subsubsection{Background}

\subsection{Limitations}
With all of the improvements seen in the past few years, however, the
synthesis community must contend with the basic limitations that the
technique will likely always have to address.  These can be split into two
main classes.  The first set of limitations has to do with the ability to
verify a program as correct or not.  Foremost among these are the difficulty
of writing accurate specifications, the unwillingness of programmers to
write these specifications, and the difficulty, in environments lacking
specifications, of determining what is the intended outcome.  These
fundamental limitations must be at the forefront of the mind of everyone who
works in this field.  They are also limiting factors to how widely these
techniques can be adopted since any critical program is unlikely to use
synthesized programs if the results they produce cannot be veified.

@@

\subsection{Approach}
Before continuing on to an explanation of our technique, there were several
assumptions and restrictions that we made in our project in order to
decrease the complexity of the programs with which we were dealing.  One of
the largest restrictions was the fact that we only tested programs that
didn't contain any type of looping structure.  These loops would have been
problematic for the following reasons.  First, since we generated symbolic
traces via symbolic execution, we would have encountered the same problem
all symbolic execution techniques do.  When encountering loops that are
either unbounded or require a large number of iterations to traverse, we
would have to unroll these loops to a certain finite depth.  This would
result in an approximation of the overall program's behavior, resulting in
an approximate fix.  This lack of certainly about the value of a fix would
be unacceptable for safety critical programs and is something that must be
dealt with before program synthesis can ever be applied in an unsupervised
setting.  Secondly, the loops would cause a problem due to the recursive
properties that would be required of a fix.  Not only would the fix be
required to fix a series of traces, the fix would be required to not
interfere with itself for a variable number of times in order to solve the
unrolled loops.  Consider for example, a loop with an error that is unrolled
2 times.  A fix would have to fix the first trace.  It would also have to be
able to fix the second trace by combining with itself in a constructive way.
This singular fix takes away the concept, seen in several papers, of
incremental improvement.  This would make it almost impossible to fix
multiple bugs that occur within a single loop, as well as more difficult to
solve a single bug.

@@

\subsubsection{Trace Generation}
The first step in synthesizing a program is generating traces that describe
the program's behavior.  Some of the techniques we read about used concrete
traces, generated from test suites.  Since we assumed that the program would
be accompanied by a comprehensive set of specs, we thought we would achieve
a much more complete understanding of the behavior of the program with
symbolic traces.  Therefore, we created a listener that would output the
complete set of instructions encountered along each symbolic path.  These
paths can be though of as being representative of an entire set of concrete
traces.  Therefore the amount of information that is captured with these
traces is much larger, resulting in a more precise analysis; one that would
be provably correct at the end.  Each of these traces is captured in SSA
form, which makes manipulating them to put into the SAT solver much easier.

\subsubsection{Trace Differentiation}
In this step of our program, we add pre and post conditions to the traces
that were generated in the previous step and calculate which traces followed
a path of execution resulting in satisfying the post condition, and which do
not.  This is a two-step process.  First, since we didn't included the
preconditions in our listener, we must remove any traces that wouldn't be
possible to follow.  We do this by querying the solver PRE AND TRACE.  This
query is akin to asking the solver, is it possible to start with values that
satisfy the precondition and execute every single instruction in the trace
without encountering any contradictions.  For example, if we generated the
trace (x1<0, x2=x1+1) from our original exploration, but the precondition
was x>0, then this trace would be completely disregarded from consideration
since its behavior is undefined.  For the traces that satisfy the
pre-conditions, there is still one more test to do.  We now must see if it
is possible to start out in the pre-conditions, execute every statement in a
trace, and end up in a space that is not considered acceptable, ie. end up
in a state that is not described by the post condition.  To do this, we ask
the solver an existential question; whether it is possible to end up in an
area that is not the post condition.  For all those traces where this is not
possible (the solver returns UNSAT), we know that these traces are safe and
all possible values that satisfy their constraints will result in satisfying
our post conditions.  For those traces where the solver returns SAT, this
means that exists at least one value that can end up outside the
post-conditions.  Therefore the entire trace is marked as problematic.  The
traces which pass the initial filter are then passed to the next stage of
the synthesis process.  Those where it is possible to end up in a space that
is not the post condition being marked ``bad,'' the other being marked
``good.''

\subsubsection{Fault Localization}

\subsubsection{Pre and Post Condition Generation}
Once the bug has been found, we now go through the process of isolating the
traces that are associated with the fault.  This is a two-part process.
First, we slice all traces, both good and bad, that go through the buggy
line so that anything at or below the bug no longer exists.  The reason for
this is that we must account for all possible ways to approach an error
within the program, but also do not want the information generated after the
bug to effect our synthesis.  Following this, we again symbolically execute
the program, this time, however, only for the paths emanating from the bug.
The reason for this is due to the fact that since the bug caused an
incorrect execution, it is possible that it forced a particular trace to go
down the wrong path following the bug.  It is also possible that a path that
was considered infeasible when executing the buggy line, now may be
possible.  Therefore, we want to consider all possible flows of execution
emanating from the error.  With the completion of this step, we have a set
of top traces and a set of bottom traces that surround the error.  This
information is then passed to the final stage of our process, synthesis, in
order to repair the program.

\subsubsection{Synthesis}
This final part of our process is by far the most intensive, requiring not
only the creation of a line that will fix the program, but also verification
that this fix works for all paths that are involved with the repair.  The
program pairs each top trace with every possible bottom trace with a series
of series of OR statements.  The result look something like [(T1 AND1 HOLE
AND1 B1) OR (T1 AND1 HOLE AND1 B2) OR ( )...] AND  [(T2 AND1 HOLE AND1 B1)
OR (T2 AND1 HOLE AND1 B2) OR ( )...] AND ... The reason for this is we know
that it is possible to traverse the top traces T1, T2, ..., since we are
assuming that there are no errors within this part of the program.  We also
know they must flow through at least one of the bottom traces (Thus the
boolean AND1's in the inner queries).  What we do not know, however, is
which bottom trace they will eventually progress down.  This is the reason
that we combine the prospective traces with OR's, since only one needs to
hold true in order for the program to work.  Finally, we combine all of the
information about how the program could progress though each of these
individual top traces together with AND2's.  This is because we program must
be able to properly execute along all possible paths that lead to the hole.

@@

\subsection{Evaluation}

\subsection{Related Work}

\subsection{Future Work}

@@

\subsection{Conclusion}

\subsection{References}

>>>>

\section{Introduction}
The field of program synthesis has seen dramatic improvements over the last
few years.  The increasing power of SAT solvers has allowed researchers to
ask more sophisticated questions and to handle larger programs than they
ever have before.  In short, this is an exciting time in program synthesis.
With the increased power available, the field has progressed beyond the
merely theoretical stage, to one that is making real contributions with even
larger gains expected in the future.

@@

\subsection{Background}

\section{Limitations}
With all of the improvements seen in the past few years, however, the
synthesis community must contend with the basic limitations that the
technique will likely always have to address.  These can be split into two
main classes.  The first set of limitations has to do with the ability to
verify a program as correct or not.  Foremost among these are the difficulty
of writing accurate specifications, the unwillingness of programmers to
write these specifications, and the difficulty, in environments lacking
specifications, of determining what is the intended outcome.  These
fundamental limitations must be at the forefront of the mind of everyone who
works in this field.  They are also limiting factors to how widely these
techniques can be adopted since any critical program is unlikely to use
synthesized programs if the results they produce cannot be veified.

@@

\section{Approach}
Before continuing on to an explanation of our technique, there were several
assumptions and restrictions that we made in our project in order to
decrease the complexity of the programs with which we were dealing.  One of
the largest restrictions was the fact that we only tested programs that
didn't contain any type of looping structure.  These loops would have been
problematic for the following reasons.  First, since we generated symbolic
traces via symbolic execution, we would have encountered the same problem
all symbolic execution techniques do.  When encountering loops that are
either unbounded or require a large number of iterations to traverse, we
would have to unroll these loops to a certain finite depth.  This would
result in an approximation of the overall program's behavior, resulting in
an approximate fix.  This lack of certainly about the value of a fix would
be unacceptable for safety critical programs and is something that must be
dealt with before program synthesis can ever be applied in an unsupervised
setting.  Secondly, the loops would cause a problem due to the recursive
properties that would be required of a fix.  Not only would the fix be
required to fix a series of traces, the fix would be required to not
interfere with itself for a variable number of times in order to solve the
unrolled loops.  Consider for example, a loop with an error that is unrolled
2 times.  A fix would have to fix the first trace.  It would also have to be
able to fix the second trace by combining with itself in a constructive way.
This singular fix takes away the concept, seen in several papers, of
incremental improvement.  This would make it almost impossible to fix
multiple bugs that occur within a single loop, as well as more difficult to
solve a single bug.

@@

\subsection{Trace Generation}
The first step in synthesizing a program is generating traces that describe
the program's behavior.  Some of the techniques we read about used concrete
traces, generated from test suites.  Since we assumed that the program would
be accompanied by a comprehensive set of specs, we thought we would achieve
a much more complete understanding of the behavior of the program with
symbolic traces.  Therefore, we created a listener that would output the
complete set of instructions encountered along each symbolic path.  These
paths can be though of as being representative of an entire set of concrete
traces.  Therefore the amount of information that is captured with these
traces is much larger, resulting in a more precise analysis; one that would
be provably correct at the end.  Each of these traces is captured in SSA
form, which makes manipulating them to put into the SAT solver much easier.

\subsection{Trace Differentiation}
In this step of our program, we add pre and post conditions to the traces
that were generated in the previous step and calculate which traces followed
a path of execution resulting in satisfying the post condition, and which do
not.  This is a two-step process.  First, since we didn't included the
preconditions in our listener, we must remove any traces that wouldn't be
possible to follow.  We do this by querying the solver PRE AND TRACE.  This
query is akin to asking the solver, is it possible to start with values that
satisfy the precondition and execute every single instruction in the trace
without encountering any contradictions.  For example, if we generated the
trace (x1<0, x2=x1+1) from our original exploration, but the precondition
was x>0, then this trace would be completely disregarded from consideration
since its behavior is undefined.  For the traces that satisfy the
pre-conditions, there is still one more test to do.  We now must see if it
is possible to start out in the pre-conditions, execute every statement in a
trace, and end up in a space that is not considered acceptable, ie. end up
in a state that is not described by the post condition.  To do this, we ask
the solver an existential question; whether it is possible to end up in an
area that is not the post condition.  For all those traces where this is not
possible (the solver returns UNSAT), we know that these traces are safe and
all possible values that satisfy their constraints will result in satisfying
our post conditions.  For those traces where the solver returns SAT, this
means that exists at least one value that can end up outside the
post-conditions.  Therefore the entire trace is marked as problematic.  The
traces which pass the initial filter are then passed to the next stage of
the synthesis process.  Those where it is possible to end up in a space that
is not the post condition being marked \emph{bad}, the other being marked
\emph{good}.

\subsection{Fault Localization}

\subsection{Pre and Post Condition Generation}
Once the bug has been found, we now go through the process of isolating the
traces that are associated with the fault.  This is a two-part process.
First, we slice all traces, both good and bad, that go through the buggy
line so that anything at or below the bug no longer exists.  The reason for
this is that we must account for all possible ways to approach an error
within the program, but also do not want the information generated after the
bug to effect our synthesis.  Following this, we again symbolically execute
the program, this time, however, only for the paths emanating from the bug.
The reason for this is due to the fact that since the bug caused an
incorrect execution, it is possible that it forced a particular trace to go
down the wrong path following the bug.  It is also possible that a path that
was considered infeasible when executing the buggy line, now may be
possible.  Therefore, we want to consider all possible flows of execution
emanating from the error.  With the completion of this step, we have a set
of top traces and a set of bottom traces that surround the error.  This
information is then passed to the final stage of our process, synthesis, in
order to repair the program.

\subsection{Synthesis}
This final part of our process is by far the most intensive, requiring not
only the creation of a line that will fix the program, but also verification
that this fix works for all paths that are involved with the repair.  The
program pairs each top trace with every possible bottom trace with a series
of series of OR statements.  The result look something like [(T1 AND1 HOLE
AND1 B1) OR (T1 AND1 HOLE AND1 B2) OR ( )...] AND  [(T2 AND1 HOLE AND1 B1)
OR (T2 AND1 HOLE AND1 B2) OR ( )...] AND ... The reason for this is we know
that it is possible to traverse the top traces T1, T2, ..., since we are
assuming that there are no errors within this part of the program.  We also
know they must flow through at least one of the bottom traces (Thus the
boolean AND1's in the inner queries).  What we do not know, however, is
which bottom trace they will eventually progress down.  This is the reason
that we combine the prospective traces with OR's, since only one needs to
hold true in order for the program to work.  Finally, we combine all of the
information about how the program could progress though each of these
individual top traces together with AND2's.  This is because we program must
be able to properly execute along all possible paths that lead to the hole.

@@

\section{Evaluation}

\section{Related Work}

\section{Future Work}

@@

\section{Conclusion}

\section{References}
